---
title: "Lab3"
author: "Patty Park"
date: "2024-04-17"
output: html_document
---
### Assignment Lab 3:

Due next week: April 23 at 11:59PM

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

#load packages
library(quanteda)
library(tm)
library(topicmodels)
library(ldatuning)
library(tidyverse)
library(tidytext)
library(reshape2)
```


For this assignment you'll use the article data you downloaded from Nexis Uni in Week 2.

```{r}
#load in data
tiktok_ca <- 
```


1.  Create a corpus from your articles.



2.  Clean the data as appropriate.

3.  Run three models (i.e. with 3 values of k) and select the overall best value for k (the number of topics) - include some justification for your selection: theory, FindTopicsNumber() optimization metrics, interpretability, LDAvis. Select the best single value of k.

4.  Plot the top terms in each topic and the distribution of topics across a sample of the documents (constrained by what looks good in the plot).

5.  Take a stab at interpreting the resulting topics. What are the key themes discussed in the articles in your data base?
